Supermartingales and Termination Analysis of Statistical PCF

### 2020-09-09 probabilisticProgrammingMartingales.tex
Added a comment on the "dependent-type style" representation of elements of \biguplus_{M \in \Lambda_0} I^{L_s(M)}

Unlike in the purely call-by-value case, the version of the reduction relation that takes into account samples is still a general relation rather than a function, so it is denoted ``$\Rightarrow$'' instead of ``$\red$'', and it relates $\biguplus_{M \in \Lambda_0} I^{L_s(M)}$ to itself.
\begin{align*}
& (M,s) \Rightarrow (N,s \circ i(M \to N)) \text{ if either} \\
& \qquad \text{$M \to N$ and the redex is not $\tsample$, or} \\
& \qquad \text{$M \mid P = \tsample$, $N = M[\underline{s(Sk(M),P)}/P]$ and $\lambda$ does not occur after $@_2$ or $\tY$ in $P$}
\end{align*}

@@

First consider the case that $P_1$ and $P_2$ are disjoint. Let $M_1 = M[X_1/P_1]$ and similarly for $X_2$, then let $M' = M[X_1/P_1][X_2/P_2]$. As the positions are disjoint, the substitutions commute and both $M_1$ and $M_2$ reduce (with $\to$) to $M'$. 
Let $s' = s \circ i(M \to M_1) \circ i(M_1 \to M')$. 
The injection $i(M \to M_1) \circ i(M_1 \to M')$ consists of \changed[lo]{prepending} $M \to M_1 \to M'$ to each reduction sequence, but by case 4.a of $\sim$, this is equivalent to \changed[lo]{prepending} $M \to M_2 \to M'$, which is $i(M \to M_2) \circ i(M_2 \to M')$. 
In the case that the reduction $M \to M_2$ isn't a $\tsample$-reduction, this is enough to establish that $(M_1, s_1) \Rightarrow (M', s')$, and vice-versa. 
If it is $\tsample$ though, in order for it to be the case that $(M_1, s_1) \Rightarrow (M', s')$, it is additionally necessary that $X_2$, the result of the reduction at $P_2$, be $s_1(M_1, P_2)$, which follows from the fact that $(M, P_2) \sim (M_1, P_2)$ by case 1. 
The case that the reduction $M \to M_1$ is a $\tsample$-reduction is similar.

>>>>

Unlike in the purely call-by-value case, the version of the reduction relation that takes into account samples is still a general relation rather than a function, so it is denoted ``$\Rightarrow$'' instead of ``$\red$'', and it relates $\biguplus_{M \in \Lambda_0} I^{L_s(M)}$ to itself.
\changed[lo]{We write an element of $\biguplus_{M \in \Lambda_0} I^{L_s(M)}$ as $(M', s)$ where the term $M' \in \Lambda^0$ identifies the component of the disjoint set, and $s \in  I^{L_s(M')}$.}
\begin{align*}
& (M,s) \Rightarrow (N,s \circ i(M \to N)) \text{ if either} \\
& \qquad \text{$M \to N$ and the redex is not $\tsample$, or} \\
& \qquad \text{$M \mid P = \tsample$, $N = M[\underline{s(Sk(M),P)}/P]$ and $\lambda$ does not occur after $@_2$ or $\tY$ in $P$}
\end{align*}

@@

First consider the case that $P_1$ and $P_2$ are disjoint. Let $M_1 = M[X_1/P_1]$ and similarly for $X_2$, then let $M' = M[X_1/P_1][X_2/P_2]$. As the positions are disjoint, the substitutions commute and both $M_1$ and $M_2$ reduce (with $\to$) to $M'$. 
Let $s' = s \circ i(M \to M_1) \circ i(M_1 \to M')$. 
The injection $i(M \to M_1) \circ i(M_1 \to M')$ consists of prepending $M \to M_1 \to M'$ to each reduction sequence, but by case 4.a of $\sim$, this is equivalent to prepending $M \to M_2 \to M'$, which is $i(M \to M_2) \circ i(M_2 \to M')$. 
In the case that the reduction $M \to M_2$ isn't a $\tsample$-reduction, this is enough to establish that $(M_1, s_1) \Rightarrow (M', s')$, and vice-versa. 
If it is $\tsample$ though, in order for it to be the case that $(M_1, s_1) \Rightarrow (M', s')$, it is additionally necessary that $X_2$, the result of the reduction at $P_2$, be $s_1(M_1, P_2)$, which follows from the fact that $(M, P_2) \sim (M_1, P_2)$ by case 1. 
The case that the reduction $M \to M_1$ is a $\tsample$-reduction is similar.
